---
Nguyễn Thanh Hòa -SE183091
Test_1
---

### 1. Install packages and library
```{r}
install.packages("RODBC")
install.packages("odbc")
install.packages("readr")
install.packages("DBI")
```

```{r}
library(RODBC)
library(odbc)
library(readr)
library(DBI)
```

### 2. Connect data Server

```{r}
server <- "localhost"
dbname <- "ChicagoPublicSchools_test1"
user <- "sa"
password <- "123456789"

# Correct the connection string format
connecttion_string <- paste0("Driver={SQL Server};",
                             "Server=", server, ";",
                             "Database=", dbname, ";",
                             "Uid=", user, ";",
                             "Pwd=", password, ";",
                             "charset=UTF-8")

# Establish the connection
con <- DBI::dbConnect(odbc::odbc(), .connection_string = connecttion_string)

# Check if the connection is successful
if (is.null(con)) {
  cat("Connection failed.\n")
} else {
  cat("Connection successful.\n")
}
```

### 3. Import data into the database

```{r}
chicago_schools <- read.csv("D:/FPT/Kihoc/fall2024/DRS301m/lab/test/ChicagoPublicSchools.csv")
dbWriteTable(con,
             "chicago_public_schools",
             chicago_schools,
             overwrite = TRUE)

print("Quá trình nhập dữ liệu đã thành công")
```

```{r}
row_count <- dbGetQuery(con,
                        "SELECT COUNT(*) AS count FROM chicago_public_schools")
cat("So luong ban ghi da dc nhap:", row_count$count, "\n")
```

```{r}
cat("Qua trinh nhap du lieu da hoan thanh\n")
```

### 4. Summarize the data

```{r}
chicago_data <- dbReadTable(con,"chicago_public_schools")
summary(chicago_data)

```

### 5. Retrieve table structure information

```{r}
columns_query <- dbGetQuery(con,
                            "SELECT School_ID
                            FROM chicago_public_schools
                            WHERE School_ID = 610281")

columns_query
```

### 6. Rename columns in the chicago_public_schools table

```{r}
dbExecute(con, "EXEC sp_rename 'chicago_public_schools.[Elementary..Middle..or.High.School]', 'School_level', 'COLUMN';")
```

### 7. Count the number of elementary schools

```{r}
query_elementary_count <- dbGetQuery(con,
                                     "SELECT COUNT(School_level) AS count 
                                     FROM chicago_public_schools 
                                     WHERE School_level = 'ES';")
query_elementary_count
```

### 8. Find the highest safety score

```{r}
query_highest_safety_score <- dbGetQuery(con,"
                                         SELECT MAX(SAFETY_SCORE) AS highest_safety_score 
                                         FROM chicago_public_schools")
query_highest_safety_score
```

### 9. Write R code to clean the columns names of a data frame called chicago_schools.

```{r}
clean_column_names <- function(names) {
  names <- gsub("+", "", names)
  names <- gsub("[._]+$", "", names)
  names <- gsub("\\.", "", names)
  return(names)
}
```

### 10. Check for missing data

```{r}
missing_data <- colSums(is.na(chicago_data))
missing_data



```
