
#SE183091_Nguyễn_Thanh_Hòa

# ASSIGMENT SQL

## PACKAGES AND LIB

```{r}
install.packages("RJDBC")

```

```{r}
system("apt-get install -y default-jdk")

```

#### JDBC driver (ex: MySQL, PostgreSQL..)

```{r}
library(RJDBC)

```

```{r}
install.packages("DBI")

```

```{r}
install.packages("RSQLite")
```

```{r}
# Tải thư viện
install.packages("rJava")
library(RJDBC)
```

```{r}
# Đường dẫn đến JDBC driver
drv <- JDBC("com.ibm.db2.jcc.DB2Driver", "C:/JavaR/jdbc_sqlj/db2jcc4.jar")


```

```{r}
install.packages("RMySQL")
```

```{r}
library(RSQLite)
```

## CONNECT DATABASE

```{r}

library(RJDBC)

# Đường dẫn đến tệp jar của MySQL Connector/J
jdbc_driver <- JDBC("com.mysql.cj.jdbc.Driver", "C:/JavaR/mysql-connector-j-9.0.0/mysql-connector-j-9.0.0/mysql-connector-j-9.0.0.jar", identifier.quote="`")

# Cung cấp thông tin kết nối
hostname <- "localhost"
port <- "3306"
dbname <- "lab1_dsr"
username <- "root1"
password <- "123456789"

# Tạo chuỗi kết nối
url <- paste0("jdbc:mysql://", hostname, ":", port, "/", dbname)

# Kết nối đến cơ sở dữ liệu
conn <- dbConnect(jdbc_driver, url, username, password)

# Kiểm tra kết nối
if (!is.null(conn)) {
  print("Kết Nối Thành Công!")
} else {
  print("Kết nối Thất Bại!")
}
```

## READ DATA
```{r}
# Đọc dữ liệu từ tệp CSV
file_path <- file.choose()
crime_data <- read.csv(file_path)
# Hiển thị một vài dòng dữ liệu để kiểm tra
head(crime_data)
# Hiển thị tổng quan dữ liệu (summary)
summary(crime_data)
```

## IMPORT ChicagoCrimeData.CSV TO DATABADE (SQL)

```{r}
# Nhập dữ liệu trực tiếp vào MySQL
dbWriteTable(conn, "ChicagoCrime", crime_data, overwrite = TRUE, row.names = FALSE)
```

## TEST DATABASE (SQL)

```{r}
# Kiểm tra dữ liệu trong bảng MySQL
result <- dbGetQuery(conn, "SELECT * FROM ChicagoCrime LIMIT 10")
print(result)
```

### Problem 1
**Total number of cases**
```{r}
query <- "
SELECT COUNT(*) AS total_cases
FROM ChicagoCrime
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 2
**Total number of cases by crime type**
```{r}
query <- "
SELECT PRIMARY_TYPE AS crime_type, COUNT(*) AS total_cases
FROM ChicagoCrime
GROUP BY PRIMARY_TYPE
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 3 **Total number of cases by year**
```{r}
query <- "
SELECT YEAR, COUNT(*) AS total_cases
FROM ChicagoCrime
GROUP BY YEAR
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 4 **Number of cases without arrests (Non-Arrest)**
```{r}
query <- "
SELECT COUNT(PRIMARY_TYPE) AS total_cases_non_arrest
FROM ChicagoCrime
WHERE ARREST = 'FALSE'
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 5 **Total number of cases by community area**
```{r}
query <- "
SELECT COMMUNITY_AREA_NUMBER, COUNT(*) AS total_cases
FROM ChicagoCrime
GROUP BY COMMUNITY_AREA_NUMBER
ORDER BY COMMUNITY_AREA_NUMBER;
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 6  Number of cases with missing location information
```{r}
query <- "
SELECT COUNT(*) AS total_cases
FROM ChicagoCrime
WHERE LOCATION_DESCRIPTION IS NULL;
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 7 **How does the trend of total crime incidents change over the years? Is there an increase or decrease?**
```{r}
query <- "
SELECT YEAR(DATE), COUNT(*) AS total_cases
FROM ChicagoCrime
GROUP BY YEAR(DATE)
ORDER BY year
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 8 **Which types of crimes are increasing or decreasing over time? Is there a specific type that is particularly prevalent in recent years?**
```{r}
query <- "
SELECT YEAR(DATE), PRIMARY_TYPE , COUNT(*) AS total_crime
FROM ChicagoCrime
GROUP BY YEAR(DATE), PRIMARY_TYPE
ORDER BY YEAR, total_crime DESC
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 9 **Which areas have a higher crime trend compared to others? Does this trend change over time?**
```{r}
query <- "
SELECT BLOCK, YEAR(DATE) , COUNT(*) AS total_crime
FROM ChicagoCrime
GROUP BY BLOCK, YEAR(DATE)
ORDER BY total_crime, YEAR(DATE) ASC
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 10 **Is there a particular time of day (morning, afternoon, evening) when crimes are more likely to occur?**
```{r}
query <- "SELECT 
  CASE 
    WHEN HOUR(DATE) < 12 THEN 'Morning'
    WHEN HOUR(DATE) < 18 THEN 'Afternoon'
    ELSE 'Evening'
  END AS time_of_day,
  COUNT(*) AS total_cases
FROM ChicagoCrime
GROUP BY time_of_day
ORDER BY FIELD(time_of_day, 'Morning', 'Afternoon', 'Evening')
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 11 **Is the crime trend affected by the season of the year? Which season typically has the highest number of incidents?**
```{r}
query <- "
SELECT 
  CASE 
    WHEN MONTH(DATE) IN (12, 1, 2) THEN 'Winter'
    WHEN MONTH(DATE) IN (3, 4, 5) THEN 'Spring'
    WHEN MONTH(DATE) IN (6, 7, 8) THEN 'Summer'
    WHEN MONTH(DATE) IN (9, 10, 11) THEN 'Fall'
  END AS Season,
  COUNT(*) AS total_cases
FROM ChicagoCrime
WHERE DATE IS NOT NULL
GROUP BY Season
ORDER BY total_cases DESC
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 12 **Which type of crime is the most prevalent in the entire dataset? What is the frequency of that type over the years?**
```{r}
query <- "
SELECT 
  PRIMARY_TYPE, 
  YEAR(DATE) ,
  COUNT(*) AS FREQUENCY
FROM ChicagoCrime
GROUP BY PRIMARY_TYPE, YEAR(DATE)
ORDER BY YEAR(DATE)  ,FREQUENCY DESC
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 13 **Which locations have the highest frequency of crimes? Are there any notable patterns or hotspots?**
```{r}
query <- "
SELECT 
  LOCATION_DESCRIPTION, 
  COUNT(*) AS total_cases
FROM ChicagoCrime
GROUP BY LOCATION_DESCRIPTION
ORDER BY total_cases DESC
LIMIT 10  
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 14 **Is there a correlation between total incidents and arrest frequency? What percentage of incidents result in arrests?**
```{r}
query <- "
SELECT 
  COUNT(*) AS total_incidents,
  SUM(CASE WHEN ARREST = TRUE THEN 1 ELSE 0 END) AS total_arrests,
  SUM(CASE WHEN ARREST = TRUE THEN 1 ELSE 0 END)/COUNT(*) as total_incidents_percent
FROM ChicagoCrime
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 15 **Which community areas have the highest frequency of crimes? Does this frequency change over time?**
```{r}
query <- "
SELECT 
    COMMUNITY_AREA_NUMBER, 
    COUNT(*) AS TotalCases 
FROM 
    ChicagoCrime 
GROUP BY 
    COMMUNITY_AREA_NUMBER 
ORDER BY 
    TotalCases DESC;
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 16 **Which location descriptions have the highest frequency of crimes? Is there a difference between various location descriptions?**
```{r}
query <- "
SELECT 
    LOCATION_DESCRIPTION, 
    COUNT(*) AS TotalCases 
FROM 
    ChicagoCrime 
GROUP BY 
    LOCATION_DESCRIPTION 
ORDER BY 
    TotalCases DESC;
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

### Problem 17 **Which types of crimes are more prevalent in specific locations? Does this trend change over time?**
```{r}
query <- "
SELECT 
    LOCATION_DESCRIPTION, 
    PRIMARY_TYPE, 
    COUNT(*) AS TotalCases 
FROM 
    ChicagoCrime 
GROUP BY 
    LOCATION_DESCRIPTION, PRIMARY_TYPE 
ORDER BY 
    LOCATION_DESCRIPTION, TotalCases DESC;
"
# Thực thi truy vấn
result <- dbGetQuery(conn, query)
# In kết quả
print(result)
```

```{r}
rmarkdown::render("SE_ASSIGMENT LAB SQL01102024.Rmd", output_format = "pdf_document")
```
