edit: Mai Tuan Anh
output:
  pdf_document: default
  html_document:
    df_print: paged
---
```{r}
# Load necessary libraries
library(dplyr)

# Read the CSV file into a dataframe
chicago_schools <- read.csv("C:/Users/Admin/Documents/Zalo Received Files/ChicagoPublicSchools.csv")
```
# 1. List the top 5 schools with the highest graduation rates
```{r}
top_graduation_schools <- chicago_schools %>%
  arrange(desc(Graduation_Rate__)) # Sort by 'Graduation_Rate__' in descending order

top_5_schools <- head(top_graduation_schools, 5) # Take the top 5 schools
```

# 2. Calculate the average safety score for each school type
```{r}
average_safety_scores <- chicago_schools %>%
  group_by("Elementary, Middle, or High School") %>%  # Group by school type
  summarise(Average_Safety_Score = mean(SAFETY_SCORE, na.rm = TRUE))  # Calculate the average safety score
```
# 3. Count the number of "Healthy School" certified schools
```{r}
healthy_school_count <- chicago_schools %>%
  filter(HEALTHY_SCHOOL_CERTIFIED == "Yes") %>%  # Filter schools with "Yes" in Healthy_School
  count()  # Count the number of such schools
```
# 4. Find the school with the highest percentage of students taking Algebra
```{r}
highest_algebra_participation <- chicago_schools %>%
  filter(!is.na(Students_Taking__Algebra__)) %>%  # Remove rows with missing Algebra participation values
  arrange(desc(Students_Taking__Algebra__)) %>%  # Sort by Algebra participation percentage
  slice(1)  # Get the school with the highest Algebra participation

```
# 5. Calculate the average ACT score for high schools
```{r}
average_act_score <- chicago_schools %>%
  filter("Elementary, Middle, or High School" == "HS") %>%  # Filter high schools
  summarise(Average_ACT_Score = mean("11th Grade Average ACT (2011)", na.rm = TRUE))  # Calculate the average ACT score

```
# 6. Count the number of schools in each community area
```{r}
schools_per_community <- chicago_schools %>%
  group_by(COMMUNITY_AREA_NAME) %>%  # Group by community area
  summarise(School_Count = n())  # Count the number of schools

```
# 7. Identify the school with the highest college enrollment rate
```{r}
highest_college_enrollment <- chicago_schools %>%
  filter(!is.na(COLLEGE_ENROLLMENT)) %>%  # Remove rows with missing college enrollment rate
  arrange(desc(COLLEGE_ENROLLMENT)) %>%  # Sort by college enrollment rate
  slice(1)  # Get the school with the highest college enrollment

```
# 8. Calculate the average environment score for each Network Manager
```{r}
average_environment_score <- chicago_schools %>%
  group_by(Network_Manager) %>%  # Group by Network Manager
  summarise(Average_Environment_Score = mean(Environment_Score, na.rm = TRUE))  # Calculate the average environment score

```
# 9. Count schools achieving "Level 1" in CPS Performance Policy
```{r}
level_1_count <- chicago_schools %>%
  filter(CPS_Performance_Policy_Level == "Level 1") %>%  # Filter schools with "Level 1"
  count()  # Count the number of such schools

```
# 10. Calculate total college enrollment by community area
```{r}
college_enrollment_by_community <- chicago_schools %>%
  filter(!is.na(COLLEGE_ENROLLMENT)) %>%  # Filter rows with valid College Enrollment data
  group_by(COMMUNITY_AREA_NAME) %>%  # Group by community area
  summarise(Total_College_Enrollment = sum(COLLEGE_ENROLLMENT, na.rm = TRUE))  # Calculate total college enrollment
```










