---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
install.packages("odbc")
install.packages("readr")
install.packages("RODBC")
install.packages('DBI')
```

### 1. Install packages and set up libraries

```{r}
library(DBI)   # Để sử dụng dbWriteTable
library(odbc)  # Nếu bạn kết nối bằng ODBC
library(readr) # Để đọc file CSV
```

### 2. Connect to the data server

```{r}
server <- "localhost,1433"  # Địa chỉ máy chủ SQL Server
database <- "DSR_demo"       # Tên cơ sở dữ liệu
user <- "sa"                 # Tên người dùng
password <- "kiet123654"    # Mật khẩu

# Tạo chuỗi kết nối
connection_string <- paste0("Driver={SQL Server};",
                            "Server=", server, ";",
                            "Database=", database, ";",
                            "Uid=", user, ";",
                            "Pwd=", password, ";",
                            "charset=UTF-8")  # Tạo kết nối hỗ trợ Unicode
```

```{r}
con <- dbConnect(odbc::odbc(), .connection_string = connection_string)

# Kiểm tra kết nối
if (is.null(con)) {
  cat("Kết nối không thành công.\n")
} else {
  cat("Kết nối thành công đến cơ sở dữ liệu ", database, ".\n")
}
```

### 3. Import data into the database

```{r}
chicago_data <- read.csv("ChicagoPublicSchool.csv", stringsAsFactors = FALSE)

dbWriteTable(con,
             name="ChiPS",
             value=chicago_data,
             overwrite=TRUE)
```

```{r}
# Kiểm tra số lượng bản ghi đã được nhập
row_count <- dbGetQuery(con, "SELECT COUNT(*) AS count 
                              FROM ChiPS")
cat("Số lượng bản ghi đã được nhập:", row_count$count, "\n")

# In ra thông báo khi quá trình nhập dữ liệu hoàn thành
cat("Quá trình nhập dữ liệu đã hoàn thành thành công!\n")
```

### 4. Summarize the data

```{r}
chicago_data <- dbReadTable(con, "ChiPS")
summary(chicago_data)
```

### 5. Retrieve table structure information

```{r}
columns_query <- dbGetQuery(con, "
                            SELECT School_ID
                            FROM ChiPS
                            WHERE School_ID = 610281")
print(columns_query)
```

### 6. Rename columns in the chicago_public_schools table

```{r}
dbExecute(con, "
          EXEC sp_rename 'ChiPS.[Elementary_Middle_or_High_School]', 'School_level', 'COLUMN';
          ")
```

### 7. Count the number of elementary schools

```{r}
query_elementary_count <- dbGetQuery(con,"
                                     SELECT COUNT(SChool_level) AS count
                                     FROM ChiPS
                                     WHERE School_level = 'ES';
                                     ")
print(query_elementary_count)
```

### 8. Find the highest safety score

```{r}
query_highest_safety_score <- dbGetQuery(con, "
                                         SELECT MAX(SAFETY_SCORE) AS highest_safety_score 
                                         FROM ChiPS
                                         ") 

print(query_highest_safety_score)
```

### 9. Write R code to clean the column names of a data frame called `chicago_schools`. 

```{r}
clean_column_names <-  function(names) {
  names <-  gsub("__+", "_", names)
  
  names <- gsub("[_\\.]+$", "", names)
  
  names <- gsub("\\.", "_", names)
  
  return(names)
}
```

### 10. Check for missing data

```{r}
missing_data <-  colSums(is.na(chicago_data))
print(missing_data)
```
